---
title: "Cyagersz_6"
author: "Cole Yagersz"
output: html_document
---

## Introduction
The purpose of this assignment is to solve a linear programming problem using the Transportation Model.

```{r}
library("lpSolveAPI")   # loads the lp solve library
```

```{r}
lprec <- make.lp(0,6)   # make an lp object with 6 decision variables and 0 constraints

set.objfn(lprec,c(622,614,630,641,645,649))   #create the objective function

lp.control(lprec,sense='min')

# Plant capacity constraints
add.constraint(lprec, c(1, 1, 1, 0, 0, 0), "<=", 100)   # Plant A
add.constraint(lprec, c(0, 0, 0, 1, 1, 1), "<=", 120)  # Plant B

# Storage space constraints
add.constraint(lprec, c(1, 0, 0, 1, 0, 0), "=", 80)  # Warehouse 1 demand
add.constraint(lprec, c(0, 1, 0, 0, 1, 0), "=", 60)  # Warehouse 2 demand
add.constraint(lprec, c(0, 0, 1, 0, 0, 1), "=", 70)  # Warehouse 3 demand



# Set bounds (non-negative)
set.bounds(lprec, lower = rep(0, 6))

write.lp(lprec,filename = "Transportation.lp", type ="lp")
lprec

# Solve and output results
solve(lprec)                 
get.objective(lprec)      # get objective value
get.variables(lprec)      # get values of decision variables
get.constraints(lprec)    # get constraints RHS values
```

Solving with creating my own lp file
```{r}
x <-read.lp("HeartStart.lp")   # Creates an lp object from reading file assign to x
x                           # displays x
```
Solve the lp model
```{r}
solve(x)
get.objective(x)     # get objective value
get.variables(x)     # get values of decision variables
get.constraints(x)   # get constraints RHS values
```

## Interpretation of Results

The transportation model for Heart Start’s AED production and distribution was solved to minimize the combined production and shipping costs while satisfying plant capacities and warehouse demands.

Optimal Total Cost:
The minimum total cost to produce and ship AEDs from the two plants to the three warehouses is $132,790. This cost includes both production and shipping expenses.

Production and Shipping Quantities:
The optimal shipment plan from plants to warehouses is:

Plant → Warehouse	Warehouse 1	Warehouse 2	Warehouse 3	Total Units Produced

Plant A	              0	         60	          40	           100

Plant B	              80	        0	          30	           110

Capacity and Demand Constraints:
Plant A produces and ships 100 units, fully utilizing its monthly production capacity.
Plant B produces and ships 110 units, which is within its capacity of 120 units.

Each warehouse’s monthly demand is fully met:
Warehouse 1: 80 units from plant B
Warehouse 2: 60 units from plant A
Warehouse 3: 70 units /40 from plant A and 30 from plant B

Interpretation of Shipment Plan:
Plant A focuses on shipping to Warehouses 2 and 3 likely because it is cheaper for plant A to produce and ship them.
Plant B ships to Warehouse 1 and some to Warehouse 3
No shipments are made from Plant A to Warehouse 1 or from Plant B to Warehouse 2, Showing that these are not as cost effective.

Summary:
The solution shows a cost-efficient distribution strategy that meets all demand and capacity constraints while minimizing overall costs. The company should fully utilize Plant A's capacity by supplying Warehouses 2 and 3, and use Plant B primarily for Warehouse 1 and part of Warehouse 3 demand.